---
title: "nalysis of MEA data and plotting - using output from matlab"
output: html_notebook
---


```{r}
library(ggplot2)
# library(RColorBrewer)
library(tidyverse)
library(ggpubr)
```


```{r}

# load data 
# in excel pre generate a file with only including well information / R error when duplicated row names 

setwd("//atlas.uni.lux/users/jennifer.modamio/DVB_group/2021/MEA/New analysis_removing outliers/3rdPlate_perElectrode_noOutliers/")


# NOTE: before and during loading, re-check the index of the rows. it is variable in some files (not fix index between extracted files)

 d15 <- read.csv("T0A_JM_15dayDiff_BrainPhys_20210211_3rd_cond_compiler.csv", header = T, skip = 156, sep=",",   nrows = 26)
 n <- d15$Measurement
 d15t <- as.data.frame(t(d15[,-1]))   
colnames(d15t) <- n  
 m <- rownames(d15t) 
 d15t %>% mutate(wells = rownames(d15t)) %>% mutate(timePoint = "day15") %>% mutate(day = 15) %>% mutate(experiment = "exp3") ->  d15t
 m -> rownames(d15t)

 d30 <- read.csv("T1A_JM_30dayDiff_BrainPhys_20210225_3rd_cond_compiler.csv", header = T, skip = 156, sep=",",     nrows = 26)
 n <- d30$Measurement
 d30t <- as.data.frame(t(d30[,-1])) 
 colnames(d30t) <- n  
 m <- rownames(d30t) 
 d30t %>% mutate(wells = rownames(d30t)) %>% mutate(timePoint = "day30") %>% mutate(day = 30)  %>% mutate(experiment = "exp3") ->   d30t
 m -> rownames(d30t)
 
 d70 <- read.csv("T2A_JM_70dayDiff_BrainPhys_3st_20210408_3rd_cond_compiler.csv", header = T,skip = 156, sep=",",   nrows = 26)
 n <- d70$Measurement
 d70t <- as.data.frame(t(d70[,-1]))  
  colnames(d70t) <- n  
   m <- rownames(d70t) 
 d70t %>% mutate(wells = rownames(d70t)) %>% mutate(timePoint = "day70") %>% mutate(day = 70) %>% mutate(experiment = "exp3") ->  d70t
 m -> rownames(d70t)
 
 d90 <- read.csv("T3A_JM_90-100dayDiff_BrainPhys_3rd_20210505_3rd_cond_compiler.csv", header = T, skip = 156, sep=",",   nrows = 26)
 n <- d90$Measurement 
 d90t <- as.data.frame(t(d90[,-1]))   
 colnames(d90t) <- n  
  m <- rownames(d90t) 
 d90t %>% mutate(wells = rownames(d90t)) %>% mutate(timePoint = "day90") %>% mutate(day = 90) %>% mutate(experiment = "exp3") ->  d90t
 m -> rownames(d90t)
 
  d120 <- read.csv("T4A_JM_120_dayDiff_BrainPhys_3rd_20210525_3rd_cond_compiler.csv", header = T, skip = 156, sep=",",   nrows = 26)
 n <- d120$Measurement 
 d120t <- as.data.frame(t(d120[,-1]))   
 colnames(d120t) <- n  
  m <- rownames(d120t) 
 d120t %>% mutate(wells = rownames(d120t)) %>% mutate(timePoint = "day120") %>% mutate(day = 120) %>% mutate(experiment = "exp3") ->  d120t
 m -> rownames(d120t)
 
 
 setwd("//atlas.uni.lux/users/jennifer.modamio/DVB_group/2021/MEA/New analysis_removing outliers/3rdPlate_perElectrode_noOutliers/")
 
 
 
 
 
```

# merge data. obtain sumary data file (over threshold and without threshold) 

```{r}

# cahnge folder for results 
#setwd("//atlas.uni.lux/users/jennifer.modamio/DVB_group/2021/MEA/New analysis_removing outliers/3rdPlate_perElectrode_noOutliers/no_ko-snca/")

setwd("//atlas.uni.lux/users/jennifer.modamio/DVB_group/2021/MEA/New analysis_removing outliers/3rdPlate_perElectrode_noOutliers/")

 ## merge all files form different time points 
dataWell <- rbind(d30t,d15t,d70t,d90t,d120t)
head(dataWell)

# add conditions to wells 


colnames(dataWell)<-c( 'NumberOfSpikes', 'MeanFiringRate', 'ISI', 'NumberOfBurst','BurstDuration_avrg', 'BurstDuration_std', 'NumberSpikesPerBurst_avg', 'NumberSpikesPerBurst_std', 'meanISI_burst_avg','meanISI_burst_std','medianISI_burst_avg','medianISI_burst_std','InterBurstInterval_avg', 'InterBurstInterval_std', 'BurstFrequency', 'IBI_CoefficientOfVariation', 'Normalized_Duration_IQR','BurstPercentage', 'condition','wells', 'timepoint', 'day','experiment')

# create colum for removal of outliers (per condition & timepoint)
dataWell %>% mutate(out = paste(dataWell$condition, dataWell$timepoint)) ->  dataWell

# convert into numeric all data of interest fro plotting (needed because data was as character) 
dataWell$NumberOfSpikes <- as.numeric(as.character(dataWell$NumberOfSpikes))
dataWell$MeanFiringRate <- as.numeric(as.character(dataWell$MeanFiringRate))
dataWell$ISI <- as.numeric(as.character(dataWell$ISI))
dataWell$NumberOfBurst <- as.numeric(as.character(dataWell$NumberOfBurst))
dataWell$BurstDuration_avrg <- as.numeric(as.character(dataWell$BurstDuration_avrg))
dataWell$BurstDuration_std <- as.numeric(as.character(dataWell$BurstDuration_std))
dataWell$NumberSpikesPerBurst_avg <- as.numeric(as.character(dataWell$NumberSpikesPerBurst_avg))
dataWell$NumberSpikesPerBurst_std <- as.numeric(as.character(dataWell$NumberSpikesPerBurst_std))
dataWell$meanISI_burst_avg <- as.numeric(as.character(dataWell$meanISI_burst_avg))
dataWell$meanISI_burst_std <- as.numeric(as.character(dataWell$meanISI_burst_std))
dataWell$medianISI_burst_std <- as.numeric(as.character(dataWell$medianISI_burst_std))
dataWell$medianISI_burst_avg <- as.numeric(as.character(dataWell$medianISI_burst_avg))
dataWell$InterBurstInterval_avg <- as.numeric(as.character(dataWell$InterBurstInterval_avg))
dataWell$InterBurstInterval_std <- as.numeric(as.character(dataWell$InterBurstInterval_std))
dataWell$BurstFrequency <- as.numeric(as.character(dataWell$BurstFrequency))
dataWell$IBI_CoefficientOfVariation <- as.numeric(as.character(dataWell$IBI_CoefficientOfVariation))
dataWell$Normalized_Duration_IQR <- as.numeric(as.character(dataWell$Normalized_Duration_IQR))
dataWell$BurstPercentage <- as.numeric(as.character(dataWell$BurstPercentage))
dataWell$day <- as.numeric(as.character(dataWell$day))


# obtain data only with bursting information for extracting into CSV and plotting in graphad (needed to take into consideration SD already calculated by software)
dataWell_burst <- dataWell[dataWell$NumberOfBurst != 0, ] #remove wells with no information for bursting 

dataWell_5spike <- dataWell[dataWell$NumberOfSpikes >= 5, ]
dataWell_5spike_burst <- dataWell_burst[dataWell_burst$NumberOfSpikes >= 5, ]


#saving info: 

  #save summary file in CSV
  write.csv(dataWell,file="dataWell_allInfo.csv")
  write.csv(dataWell_5spike,file="dataWell_5spike.csv")
  # save information reagrding average and std
  write.csv(dataWell_burst,file="dataWell_burst_allInfo.csv") 
  write.csv(dataWell_5spike_burst,file="dataWell_5spike_burst.csv")


```


# check normality (shapiro.test) 

# Remove outliers based on 25th & 75th percentiles (for all wells based on a single feature)

```{r} 
# NOTE!! if zeros are kept, may infomation will be extracted based on a simple value due to the ceros being the mean !! 
# transfrom 0 into NAs 


dataWell_noZero <- replace(dataWell_5spike, dataWell_5spike == 0, NA)

 groupOut <- unique(dataWell_noZero$out)
 
 #groupOut <- c("WT day30" ,"3xSNCA day30" ,"KO-SNCA day30" ,"WT day15", "3xSNCA day15" , "KO-SNCA day15", "WT day70" ,"3xSNCA day70","KO-SNCA day70" ,"WT day90","3xSNCA day90","KO-SNCA day90")
 
 
 index <- data.frame()
 index2 <- data.frame()
 index3 <- data.frame()
 index4 <- data.frame()

for (i in 1:length(groupOut))
{
  datatest <- dataWell_noZero %>% 
  filter(out %in% groupOut[i]) # go for each line and extract rows for that line 
  #exprs <- datatest[2:24] # extract the variables within that line 
  feature_names <- colnames(datatest[1:18])  # create vector with names of variables 
  for (j in 1:19) # for (j in 2:length(exprs))
  {
    out <- boxplot.stats(datatest[,j])$out
    index2 <- data.frame()
    for (k in 1:length(out))
    {
    index <- data.frame()
    index <- datatest[datatest[,j]==out[k],]# sum up all index of all data so it can be removed all lies afterwards 
    index2 <- rbind(index2,index)
    rm(index)
    }
    index2 %>% mutate(cond_outlier = feature_names[j]) %>% mutate(group_tested = groupOut[i]) -> index3
    index4 <- rbind(index4,index3)
    rm(out, index2, index3)
  }
} 


# remove data from extra "NA"

 summary_out <- unique(index4[1:24])
 summary_in <- setdiff(dataWell_noZero, summary_out)
 
# Remove an entire row if it has >22 NAs (for example >50% of your features)
count_na <- function(x)sum(is.na(x))
data_noOut_well <- summary_in %>%
  dplyr::mutate(count_na = apply(., 1, count_na))
data_noOut_well <-summary_in[!(data_noOut_well$count_na >= 19),] # remove wells containing no info 


# remove the data containinf outlayers from main dataset * a outlayer in one single feature will cause the removal of full well data 

dataWell_burst_NoOut_well <- data_noOut_well[data_noOut_well$NumberOfBurst != 0, ] #remove wells with no information for bursting

#saving info: 
# files without outliers were generated from file No zeros - reason: too many zeros cause exclusion of real values as outliers 
# _well > meaningn all well information was removed for all variable when at least one of the variables presented one outlier -> the well is not consider any longer 

  write.csv(index4,file="index4_info_outliers.csv") # wells excluded, reason for exclusion and group 
  write.csv(summary_out,file="summary_out_outliers.csv") # summary of wells excluded (no accounting reason)
  write.csv(dataWell_noZero,file="dataWell_noZero.csv") # data well _ full information _ no zeros  
  write.csv(dataWell_burst_NoOut_well,file="dataWell_burst_NoOut_well.csv") # infor for bursting (graphad) _no outliers *one oulier value excludes the whole well 
  write.csv(data_noOut_well,file="data_noOut_well.csv") # infor - no outliers *one oulier value excludes the whole well 
  
  
```

# Remove outliers based on 25th & 75th percentiles (for only the specific feature under study_ introducetion NAs)
# thes rest of the information for the full well remains 


```{r} 
# NOTE!! if zeros are kept, may infomation will be extracted based on a simple value due to the ceros being the mean !! 
# transfrom 0 into NAs 


#dataWell_noZero <- replace(dataWell, dataWell == 0, NA)
datatest_NA <- data.frame()


for (i in 1:length(groupOut))
{
  datatest <- dataWell_noZero %>% 
  filter(out %in% groupOut[i]) # go for each line and extract rows for that line 
  for (j in 1:19) # for (j in 2:length(exprs))
  {
    out <- boxplot.stats(datatest[,j])$out
    for (k in 1:length(out))
    {
    datatest[,j][datatest[,j] == out[k]] <- NA
    }
    rm(out)
  }
  datatest_NA <- rbind(datatest_NA,datatest)
  rm(datatest)
} 


#saving info: 
# files without outliers were generated from file No zeros - reason: too many zeros cause exclusion of real values as outliers 
# _well > meaningn all well information was removed for all variable when at least one of the variables presented one outlier -> the well is not consider any longer 

# Remove an entire row if it has >22 NAs (for example >50% of your features)
count_na <- function(x)sum(is.na(x))
data_noOut_value <- datatest_NA %>%
  dplyr::mutate(count_na = apply(., 1, count_na))
data_noOut_value <-datatest_NA[!(data_noOut_value$count_na >= 19),] # remove wells containing no info 

dataWell_burst_NoOut_value <- subset(data_noOut_value, !is.na(NumberOfBurst)) #remove wells with no information for bursting

  write.csv(dataWell_burst_NoOut_value,file="dataWell_burst_NoOut_value.csv") # infor for bursting (graphad) _ only the specific outlier was removed. well information for other variables remain  
  write.csv(data_noOut_value,file="data_noOut_value.csv") # infor - no outliers for specific variable -well info remains 
  
```

```{r warning=FALSE}

# removing data fro KO-SNCA line 
data_noOut_value_KO = filter(data_noOut_value, !(condition %in% "KO-SNCA"))

# cahnge folder for results 
setwd("//atlas.uni.lux/users/jennifer.modamio/DVB_group/2021/MEA/New analysis_removing outliers/3rdPlate_perElectrode_noOutliers/no_ko-snca/")

write.csv(data_noOut_value_KO,file="data_noOut_value_KO.csv") # infor - no outliers for specific variable -well info remains 
 
```


# plotting without SD already calculated ("NumberOfSpikes" "MeanFiringRate"  "ISI"  "NumberOfBurst"  "NumberBurstingElectrodes")


```{r warning=FALSE}
# datasets to test - remember to modify the saving name PDF 
# dataWell
# dataWell_noZero
# data_noOut_well
# data_noOut_value 


head(data_noOut_value_KO)
feature_names <- colnames(data_noOut_value_KO[,c(1:4,15:18)])  # extract only values (no average)
# interspike interval [ISI]
names <- c("Number Of Spikes", "Mean Firing Rate" , "Interspike Interval [ISI]"  ,"Number Of Burst" , "Burst Frequency","IBI Coefficient Of Variation", "Normalized Duration IQR"  ,  "Burst Percentage")


for (i in 1:length(feature_names)) {
data_noOut_value_KO %>%  # change dataset selected - change saving option depending on the dataset 
  pivot_longer(cols=feature_names, names_to = "feature", values_to = "value") %>%
  filter(feature %in% feature_names[i]) %>%
  ggplot( aes(x = factor(condition, level = c( 'WT','3xSNCA')), y=value) ) +
  geom_boxplot(aes(fill=fct_relevel(condition, 'WT','3xSNCA')),width=0.5,outlier.shape=NA) +
  scale_fill_manual(values= c("#008B8B","#B22222","#2171b5"),name = "condition") +
   # t-test
  #stat_compare_means(comparisons=list(c("WT", "3xSNCA")), method = "t.test", label="p.signif", label.x = 1.5)+
  
  # Mann-Whitney
  ggpubr::stat_compare_means(comparisons=list(c("WT", "3xSNCA")), method="wilcox.test", p.adjust.method="BH",label="p.signif", label.x = 1.5)+
  
    
  facet_grid(~fct_relevel(timepoint, "day15","day30", "day70", "day90", "day120"), scales="free") +
  labs(x     ="",
       y     = paste(names[i]),
       fill  = "Condition",
       title = paste(names[i])) +
  theme_bw() +
    
  theme(
    axis.line = element_line(colour = 'black', size = 0.1) ,
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 5), #angle=45),
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size=9, color="black"),
    axis.ticks.y = element_line(),
    legend.position="right",
    legend.text = element_text(size = 11, face = "bold") ,
    legend.title = element_blank(),
    legend.key.size = unit(0.7, "cm"),
    legend.key.width = unit(0.6,"cm"),
    plot.title = element_text(size = 15, hjust=0.5, vjust= 1, face = "bold"),
    plot.subtitle = element_blank(),#element_text(size = 2, hjust=0.5)
    strip.text = element_text(size=12, vjust=0.5),
    strip.background = element_rect(fill="lightgray"),
    panel.border = element_rect(fill = NA, color = "black"),
    panel.spacing.y = unit(0.8, "lines"),
    strip.switch.pad.wrap=unit(20, "lines")
  )  -> p
  t<- cowplot::ggdraw(cowplot::add_sub(p, "Wilcox-test, ***p=0.001, **p=0.01, *p=0.05",hjust=-0.5, size=9))
  print(t)

  ggsave(paste0(Sys.Date(),"_", feature_names[i], "_data_noOut_value_electrode.pdf"), plot=t) # change saving option depending on the dataset 

}

```


# plot into timeline with Standard deviation 

```{r warning=FALSE}

#"NumberOfSpikes" 
#"MeanFiringRate" 
#"ISI" 
#"NumberOfBurst" 
#"NumberBurstingElectrodes"

data_noOut_value_KO %>%
  ggplot(aes(x=day, y=NumberOfSpikes, group=condition, color=condition)) +
  stat_smooth(method='loess', formula ='y ~ x', size = 1)+
  scale_colour_manual(name="Condition", values= c("WT"= "#008B8B", "3xSNCA"="#B22222")) +
  # facet_wrap(~timepoint, scales="free", labeller = labeller(groupwrap = label_wrap_gen(10))) +
  # facet_grid(~fct_relevel(timepoint, "day15","day30",  "day70", "day90"), scales="free") +
  labs(x="Time of differentiation [days]",
        y     = "Number of Spikes",
       fill  = "Condition",
       title = "Total number of Spikes") +
   theme_bw() -> p
  t<- cowplot::ggdraw(cowplot::add_sub(p, "Locally estimated scatterplot smoothing: y ~ Time",hjust=-0.2, size=9))
  print(t)
  ggsave(paste0(Sys.Date(),"NumberOfSpikes_timeline_data_noOut_value.pdf"), plot=t)
```

```{r warning=FALSE}

#"NumberOfSpikes" 
#"MeanFiringRate" 
#"ISI" 
#"NumberOfBurst" 
#"NumberBurstingElectrodes"

data_noOut_value_KO %>%
  ggplot(aes(x=day, y=MeanFiringRate, group=condition, color=condition)) +
  stat_smooth(method='loess', formula ='y ~ x', size = 1)+
  scale_colour_manual(name="Condition", values= c("WT"= "#008B8B", "3xSNCA"="#B22222")) +
  # facet_wrap(~timepoint, scales="free", labeller = labeller(groupwrap = label_wrap_gen(10))) +
  # facet_grid(~fct_relevel(timepoint, "day15","day30",  "day70", "day90"), scales="free") +
  labs(x="Time of differentiation [days]",
        y     = "Mean Firirng rate",
       fill  = "Condition",
       title = "Mean Firirng rate") +
   theme_bw() -> p
  t<- cowplot::ggdraw(cowplot::add_sub(p, "Locally estimated scatterplot smoothing: y ~ Time",hjust=-0.2, size=9))
  print(t)
  ggsave(paste0(Sys.Date(),"MeanFiringRate_timeline_data_noOut_value.pdf"), plot=t)
```

```{r warning=FALSE}

#"NumberOfSpikes" 
#"MeanFiringRate" 
#"ISI" 
#"NumberOfBurst" 
#"NumberBurstingElectrodes"

data_noOut_value_KO %>%
  ggplot(aes(x=day, y=ISI, group=condition, color=condition)) +
  stat_smooth(method='loess', formula ='y ~ x', size = 1)+
  scale_colour_manual(name="Condition", values= c("WT"= "#008B8B", "3xSNCA"="#B22222")) +
  # facet_wrap(~timepoint, scales="free", labeller = labeller(groupwrap = label_wrap_gen(10))) +
  # facet_grid(~fct_relevel(timepoint, "day15","day30",  "day70", "day90"), scales="free") +
  labs(x="Time of differentiation [days]",
        y     = "[ISI]",
       fill  = "Condition",
       title = "interspike interval [ISI]") +
   theme_bw() -> p
  t<- cowplot::ggdraw(cowplot::add_sub(p, "Locally estimated scatterplot smoothing: y ~ Time",hjust=-0.2, size=9))
  print(t)
  ggsave(paste0(Sys.Date(),"ISI_timeline_data_noOut_value.pdf"), plot=t)
```


  
  Extract later time points 
  
```{r warning=FALSE}

# removing later time points for visualization 

data_noOut_value_time = filter(data_noOut_value_KO, !(timepoint %in% "day120"))
data_noOut_value_time = filter(data_noOut_value_time, !(timepoint %in% "day90"))
```
  
```{r warning=FALSE}

#"NumberOfSpikes" 
#"MeanFiringRate" 
#"ISI" 
#"NumberOfBurst" 
#"NumberBurstingElectrodes"

data_noOut_value_time %>%
  ggplot(aes(x=day, y=MeanFiringRate, group=condition, color=condition)) +
  stat_smooth(method='loess', formula ='y ~ x', size = 1)+
  scale_colour_manual(name="Condition", values= c("WT"= "#008B8B", "3xSNCA"="#B22222")) +
  # facet_wrap(~timepoint, scales="free", labeller = labeller(groupwrap = label_wrap_gen(10))) +
  # facet_grid(~fct_relevel(timepoint, "day15","day30",  "day70", "day90"), scales="free") +
  labs(x="Time of differentiation [days]",
        y     = "Mean Firirng rate",
       fill  = "Condition",
       title = "Mean Firirng rate") +
   theme_bw() -> p
  t<- cowplot::ggdraw(cowplot::add_sub(p, "Locally estimated scatterplot smoothing: y ~ Time",hjust=-0.2, size=9))
  print(t)
  ggsave(paste0(Sys.Date(),"MeanFiringRate_timeline_data_noOut_value_time.pdf"), plot=t)
```

```{r warning=FALSE}

#"NumberOfSpikes" 
#"MeanFiringRate" 
#"ISI" 
#"NumberOfBurst" 
#"NumberBurstingElectrodes"

data_noOut_value_time %>%
  ggplot(aes(x=day, y=NumberOfSpikes, group=condition, color=condition)) +
  stat_smooth(method='loess', formula ='y ~ x', size = 1)+
  scale_colour_manual(name="Condition", values= c("WT"= "#008B8B", "3xSNCA"="#B22222")) +
  # facet_wrap(~timepoint, scales="free", labeller = labeller(groupwrap = label_wrap_gen(10))) +
  # facet_grid(~fct_relevel(timepoint, "day15","day30",  "day70", "day90"), scales="free") +
  labs(x="Time of differentiation [days]",
        y     = "[ISI]",
       fill  = "Condition",
       title = "interspike interval [ISI]") +
   theme_bw() -> p
  t<- cowplot::ggdraw(cowplot::add_sub(p, "Locally estimated scatterplot smoothing: y ~ Time",hjust=-0.2, size=9))
  print(t)
  ggsave(paste0(Sys.Date(),"NumberOfSpikes_timeline_data_noOut_value_time.pdf"), plot=t)
```
